---
title: "Week1_Data_Type"
author: "Jinsu Han"
date: "Tuesday, February 17, 2015"
output: html_document
---
## Entering Input

At the R prompt we type expressions. The <- Symbol is the assignment operator.
```  
x <-1
print(x)
[1] 1
x
[1] 1
msg <- "hello"
```

The grammer of the language determines whether an expression is complete or not.
```
x <- ## Incomplete expression
```

The # character indicates a comment. Anything to the right of the # (including the # itself) is ignored.

## Evaluation

When a complete expression is entered at the prompt, it is evaluted and the result of the evaluated expression is returned. The result may be auto-printed.
```
x <- 5 ## nothing printed
x      ## auto-printing occurs
[1] 5
print(x)  ## explict printing
[1] 5
```

The [1] indicates that x is a vector and 5 is the first element.

## Printing
```{r}
x <- 1:20
x
```
The : operator is used to create integer sequences.

## Creating Vectors
The c() function can be used to create vectors of objects
```
x <- c(0.5,0.6)     ## numeric
x <- c(TRUE,FALSE)  ## logical
x <- c(T,F)         ## logical
x <- c("a","b","c") ## character
x <- 9:29           ## integer
x <- c(1+0i, 2+4i)  ## complex
```
Using the vector() function
```{r}
x <-vector("numeric", length=10)
x
```

## Mixing Objects
What about the following?
```
y <- c(1.7, "a")  ## character
y <- c(TRUE, 2)   ## numeric
y <- c("a", TRUE) ## character
```
When different objects are mixed in a vector, coercion occurs so that every element in the vector is of the same class.

## Explicit Coercion
Objects can be explicitly coerced from one class to another using the as.*functions, if available.
```{r}
x <-0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```

Nonsensical coercion results in NAS.
```{r}
x <-c("a","b","c")
as.numeric(x)
as.logical(x)
as.complex(x)
```

## Lists
Lists are a special type of vector that can contain elements of different classes. Lists are a very important data type in R and you should get to know them well.

```{r}
x <- list(1, "a", TRUE, 1+4i)
x
```

## Matrices
Matrices are vectors with a dimension attribute. The dimension attribute is itself an integer vector of length 2(nrow,ncol)

```{r}
m <- matrix(nrow = 2, ncol = 3)
m
dim(m)
attributes(m)
```

## Matrices(cont'd)
Matrices are constructed column-wise, so entries can be thought of starting in the "upper left" corner and running down the columns.

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
m
```
## Matrices(cont'd)
Matrices can also be created directly from vectors by adding a dimension attribute.
```{r}
m <- 1:10
m
dim(m) <- c(2,5)
m
```

## cbind-ing and rbind-ing
Matrices can be created by column-binding or row-binding with cbind() and rbind()
```{r}
x <- 1:3
y <- 10:12
cbind(x,y)
rbind(x,y)
```

## Factors
Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer has a label.  

* Factors are treated specially by modelling functions like `lm()` and `glm()`  

* Using factors with labels is better than using integers because factors are self-describing; having a variable that has values "Male" and "Female" is better than a variable that has values 1 and 2.

```{r}
x <- factor(c("yes","yes","no","yes","no"))
x
table(x)
unclass(x)
```

The order of the levels can be set using the levels argument to `factor()`. This can be important in linear modelling because the first level is used as the baseline level.

```{r}
x <- factor(c("yes","yes","no","yes","no"),levels =c("yes","no"))
x
```

## Missing Values
Missing values are devoted by `NA` or `NaN` for undefined mathematical operations.  
* `is.na()` is used to test objects if they are `NA`  
* `is.na()`is used to test for `NaN`  
* `NA` values have a class also, so there are integer `NA`, character `NA`, etc.  
* A NaN value is also `NA` but the converse is not true.

```{r}
x <- c(1,2,NA,10,3)
is.na(x)
is.nan(x)
x <- c(1,2,NaN,NA,4)
is.na(x)
is.nan(x)
```

## Data Frames
Data frames are used to store tabular data.    

* They are represented as a special type of list where every element of the list has to have the same length  

* Each element of the list can be thought of as a column and the length of each element of the list is the number of rows  

* Unlike matrices, data frames can store different classes of objects in each column (just like lists); matrices must have every element be the same class  

* Data frames also have a special attribute called `row.names`  

* Data frames are usually created by calling `read.table()` or `read.csv()`  

* Can be converted to a matrix by calling `data.matrix(x)`  

```{r}
x <- data.frame(foo = 1:4, bar = c(T,T,F,F))
x
nrow(x)
ncol(x)
```

## Names
R objects can also have names, which is very useful for writing readable code and self-describing objects.

```{r}
x <- 1:3
names(x)
names(x) <- c("foo","bar","norf")
x
names(x)
```

Lists can also have names.
```{r}
x <- list(a=1, b=2, c=3)
x
```
And matrices.
```{r}
m <- matrix(1:4, nrow = 2, ncol=2)
dimnames(m) <- list(c("a","b"),c("c","d"))
m
```

## Summary
Data Types   

* atomic classes : numeric, logical, character, integer, complex\  

* vectors, lists  

* factors  

* missing values  

* data frames  

* names  
